Relative parabolic factorizations of Chevalley groups

### 2015-05-23 bass-kolster.tex
Minor changes

It is easy to see that the proof of the above theorem is effective and gives an estimate of the total number of elementary root unipotents involved in the decomposition.
\begin{cor}
In the assumptions and notation of Theorem~\ref{thm:BassKolster} every element of $\G(\Phi,R,I)$ 
can be factored into a product of one element of $\G(\Delta_1,R,I)$ one element of $Z$ and at most $n(\Phi)$ elementary root unipotents $t_{\alpha}(s)$ of level $I$, where
$n(\rA_\ell)= 4 \ell - 1$, $n(\rB_\ell)= 8 \ell - 5$, $n(\rC_\ell)= 8 \ell - 5$, $n(\rD_\ell)= 8 \ell - 9$. \end{cor}
\begin{proof}
The estimates stated above can be obtain by careful analysis of the proof of the previous theorem.
Moreover, the estimates for $\Phi=\rA_\ell, \rC_\ell$ are immediate.
In the case $\Phi=\rD_\ell$ the proof of Theorem~\ref{thm:BassKolster} implies 
\begin{multline}\nonumber
\G(\Phi,R,I)=  \U(\Sigma_\ell,I) \cdot X_{\alpha_1}(I) \cdot \U(\Sigma_2^-\cap\Delta_1,I) \cdot X_{-\alpha\ssub{\mathrm{max}}}(I) \cdot Z  \cdot \\ \cdot \U(\Sigma_1^-,I) \cdot \U(\Sigma_1,I) \cdot \G(\Delta_1,R,I).
\end{multline}
We can present an element $g$ of $\U(\Sigma_\ell, I)$ as a product of $g_1 \in \U((\Sigma_1 \cup \Sigma_2) \cap \Sigma_\ell)$ and $g_2\in \U(\Sigma_\ell\cap\Delta_1\cap\Delta_2)$.
An examination of the extended Dynkin diagram of $\rD_\ell$ implies that $g_2$ either centralizes or normalizes all factors of the above decomposition (except the last one) and therefore can be safely moved to the right.
On the other hand, $g_1$ is a product of at most $2\ell-3$ elementary unipotents, while the width of $\U(\Sigma_1^\pm, I)$ and $\U(\Sigma_2^-\cap\Delta_1)$ in elementary unipotents does not exceed $2\ell-2$ and $2\ell-4$, respectively.
Summing up these upper bounds we obtain
$$n(\rD_\ell)\leq (2\ell-3) + 1 + (2\ell - 4) + 1 + 2\cdot (2\ell - 2) = 8\ell - 9.$$

The estimate for the case $\Phi=\rB_\ell$ can be obtained by a similar argument. \end{proof}

>>>>

It is easy to see that the proof of the above theorem is effective and gives an estimate of the total number of elementary root unipotents involved in the decomposition.
\begin{cor}
In the assumptions and notation of Theorem~\ref{thm:BassKolster} every element of $\G(\Phi,R,I)$ 
can be factored into a product of one element of $\G(\Delta_1,R,I)$ one element of $Z$ and at most $n(\Phi)$ elementary root unipotents $t_{\alpha}(s)$ of level $I$, where
$n(\rA_\ell)= 4 \ell - 1$, $n(\rB_\ell)= 8 \ell - 5$, $n(\rC_\ell)= 8 \ell - 5$, $n(\rD_\ell)= 8 \ell - 9$. \end{cor}
\begin{proof}
The above estimates can be obtained by a careful analysis of the proof of the previous theorem.
Cases $\Phi=\rA_\ell, \rC_\ell$ are immediate.
In the case $\Phi=\rD_\ell$ the proof of Theorem~\ref{thm:BassKolster} implies that
\begin{multline}\nonumber
\G(\Phi,R,I) =  \U(\Sigma_\ell,I) \cdot X_{\alpha_1}(I) \cdot \U(\Sigma_2^-\cap\Delta_1,I) \cdot X_{-\alpha\ssub{\mathrm{max}}}(I) \cdot Z  \cdot \\ \cdot \U(\Sigma_1^-,I) \cdot \U(\Sigma_1,I) \cdot \G(\Delta_1,R,I).
\end{multline}
We can present an element $g$ of $\U(\Sigma_\ell, I)$ as a product of $g_1 \in \U((\Sigma_1 \cup \Sigma_2) \cap \Sigma_\ell)$ and $g_2\in \U(\Sigma_\ell\cap\Delta_1\cap\Delta_2)$.
An examination of the extended Dynkin diagram of $\rD_\ell$ implies that $g_2$ either centralizes or normalizes all factors of the above decomposition (except the last one) and therefore can be safely moved to the right.
On the other hand, $g_1$ is a product of at most $2\ell-3$ elementary unipotents, while the width of $\U(\Sigma_1^\pm, I)$ and $\U(\Sigma_2^-\cap\Delta_1)$ in elementary unipotents does not exceed $2\ell-2$ and $2\ell-4$, respectively.
Summing up these upper bounds we obtain
$$n(\rD_\ell) = (2\ell-3) + 1 + (2\ell - 4) + 1 + 2\cdot (2\ell - 2) = 8\ell - 9.$$

The estimate in the case $\Phi=\rB_\ell$ can be obtained in a similar way. \end{proof}

### 2015-05-23 dennis-vaserstein.tex
Minor changes

Throughout the present section we denote by $\EP_s(R, I)$ the subgroup $\E(\Delta_s, R, I) \cdot \U(\Sigma_s, I)$, $1 \leq s \leq n$.
%TODO: Add Levi decomposition to preliminaries
%TODO: Introduce notation for U(S, R)
Set $\EP_s := \EP_s(R, R) = \E(\Delta_s, R) \cdot \U(\Sigma_s, R)$.

@@

From now on we can assume that the conlcusion of Lemma~\ref{lemma:Stein_reduction} holds and $\EP_s$ normalizes $A_{rs}$.
Notice that in view of Theorem~\ref{theorem:Stepanov} it suffices to show that the following two families of elements are contained in $L_{rs}$:
\begin{itemize} \item $z_{\alpha}(s, \xi)$, $s\in I$, $\xi \in R$, $\alpha\in\Sigma^-_s$;
\item $x_{\beta}(s)$, $s \in I$, $\beta \in \Phi$. \end{itemize}
Since $\EP_s \subseteq N_{rs}$ it suffices to check inclusions only for the second family of elements.
We already know that $\U^+(\Phi, I) \subseteq L_{rs}$.

>>>>

@@

From now on we can assume that the conclusion of Lemma~\ref{lemma:Stein_reduction} holds and $\EP_s$ normalizes $A_{rs}$.
Notice that in view of Theorem~\ref{theorem:Stepanov} it suffices to show that the following two families of elements are contained in $L_{rs}$:
\begin{itemize} \item $z_{\alpha}(s, \xi)$, $s\in I$, $\xi \in R$, $\alpha\in\Sigma^-_s$;
\item $x_{\beta}(s)$, $s \in I$, $\beta \in \Phi$. \end{itemize}
Since $\EP_s \subseteq N_{rs}$ it suffices to check inclusions only for the second family of elements.
We already know that $\U^+(\Phi, I) \subseteq L_{rs}$.

### 2015-05-23 paper.tex
Minor changes

The lemma is a relative version of the classical result sometimes called Chevalley---Matsumoto decomposition.
%TODO: Recheck the statement of the lemma; finish the proof.
\begin{lemma}\label{lemma:Chevalley-Matsumoto}
Let $\pi$ be the fundamental representation of $\G_{sc}(\Phi, R)$ with the highest weight $\varpi_s$.
Assume that $g\in \G_{sc}(\Phi, R, I)$ is such that $(g\cdot v^+)_{\varpi\ssub{s}}=1$, then 
$g \in \U(\Sigma_s^-, I) \cdot \G_{sc}(\Delta_s, R, I) \cdot \U(\Sigma_s^+, I)$.
\end{lemma} \begin{proof}  For $I=R$ the statement of the lemma is contained in~\cite[Theorem~1.3]{St78}). \end{proof}

@@

>>>>

The following lemma is a relative version of the classical result sometimes called Chevalley---Matsumoto decomposition.
%TODO: Recheck the statement of the lemma; finish the proof.
\begin{lemma}\label{lemma:Chevalley-Matsumoto}
Let $\pi$ be the fundamental representation of $\G_{sc}(\Phi, R)$ with the highest weight $\varpi_s$.
Assume that $g\in \G_{sc}(\Phi, R, I)$ is such that $(g\cdot v^+)_{\varpi\ssub{s}}=1$, then 
$g \in \U(\Sigma_s^-, I) \cdot \G_{sc}(\Delta_s, R, I) \cdot \U(\Sigma_s^+, I)$.
\end{lemma} \begin{proof}  For $I=R$ the statement of the lemma is contained in~\cite[Theorem~1.3]{St78}). \end{proof}

@@

Set $\EP_s(R, I) = \E(\Delta_s, R, I) \cdot \U(\Sigma_s, I)$, $1 \leq s \leq n$.
It is easy to see that $\EP_s(R, I)$ is a subgroup.
 Throughout the present section we denote by
%TODO: Add Levi decomposition to preliminaries
%TODO: Introduce notation for U(S, R)
Set $\EP_s := \EP_s(R, R) = \E(\Delta_s, R) \cdot \U(\Sigma_s, R)$.
